---
title: "DATA 607  PROJECT 4: Document Classification"
author: "Yohannes Deboch"
date: "04/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load the libraries
```{r}
library(caret)
library(MASS)
library(pROC)
```

# Read the data set
```{r}
spambase <- read.csv("spambase.data.txt", header = FALSE)
head(spambase)
```

Assign column names
```{r}
colnames(spambase) <-  c("word_freq_make", "word_freq_address", "word_freq_all", "word_freq_3d", 
    "word_freq_our", "word_freq_over", "word_freq_remove", "word_freq_internet", 
    "word_freq_order", "word_freq_mail", "word_freq_receive", "word_freq_will", 
    "word_freq_people", "word_freq_report", "word_freq_addresses", "word_freq_free", 
    "word_freq_business", "word_freq_email", "word_freq_you", "word_freq_credit", 
    "word_freq_your", "word_freq_font", "word_freq_000", "word_freq_money", 
    "word_freq_hp", "word_freq_hpl", "word_freq_george", "word_freq_650", "word_freq_lab", 
    "word_freq_labs", "word_freq_telnet", "word_freq_857", "word_freq_data", 
    "word_freq_415", "word_freq_85", "word_freq_technology", "word_freq_1999", 
    "word_freq_parts", "word_freq_pm", "word_freq_direct", "word_freq_cs", "word_freq_meeting", 
    "word_freq_original", "word_freq_project", "word_freq_re", "word_freq_edu", 
    "word_freq_table", "word_freq_conference", "char_freq_ch;", "char_freq_ch(", 
    "char_freq_ch[", "char_freq_ch!", "char_freq_ch$", "char_freq_ch#", "capital_run_length_average", 
    "capital_run_length_longest", "capital_run_length_total", "spam")
head(spambase)
```

```{r}
barplot(table(spambase$spam),main="Spam vs Non Spam")
```
In the data most of the emails are non smap.

# Split into training and test set
```{r}
# Set seed value
set.seed(123)
# Shuffle the data set in R
spambase <- spambase[sample(nrow(spambase)),]
#converst spam to factor

spambase$spam <- as.factor(spambase$spam)
index <- createDataPartition(spambase$spam, p = 0.70,list = FALSE)
training <- spambase[index,]
test <- spambase[-index,]
```


# Build model on Training Data
I've used SVM,knn and LDA for spambase classification

```{r}

# Control parameters for model
train_ctrl <- trainControl(method = "repeatedcv",number = 10, repeats = 10)

model_svm <- train(spam~., data = training, method = "svmLinear",trControl = train_ctrl)
model_lda <- train(spam~., data = training, method = "lda",trControl = train_ctrl)
model_knn <- train(spam~., data = training, method = "knn",trControl = train_ctrl)

```

# Do prediction
```{r}
# Do prediction
pred_svm <- predict(model_svm, test[,-58])
pred_lda <- predict(model_lda, test[, -58])
pred_knn <- predict(model_knn, test[,-58])


```

# Measure Accuracy on Test Data

# Confusion Matrix
```{r}
paste("SVM confusion matrix")
confusionMatrix(pred_svm,test$spam)
paste("LDA confusion matrix")
confusionMatrix(pred_lda,test$spam)
paste("KNN confuxion matrix")
confusionMatrix(pred_knn,test$spam)
```

Out of smv,lda and knn model SVM has best test performance with test accuracy 0.9217    .
# ROC Curve
```{r}

# calculate roc curve values
roc_svm <- roc(pred_svm,as.numeric(test$spam))
roc_lda <- roc(as.numeric(pred_lda),as.numeric(test$spam))
roc_knn <- roc(as.numeric(pred_knn), as.numeric(test$spam))

# Area under the curve
roc_svm$auc
roc_lda$auc
roc_knn$auc
plot(roc_svm,col="red",lty=1,main="ROC curve for 3 model")
plot(roc_lda,col="green",lty=2,add=TRUE)
plot(roc_knn,col="blue",lty=3,add=TRUE) 
legend("topright", legend=c("SVM","LDA","KNN"), col = c("red","green","blue"), lty=1:3)
```

From the ROC curve we can that SVM has the highest prediction performance.